CUBE: An Information-optimized parallel Cosmological $N$-body Algorithm

### 2017-12-17 ms_caf/ms_caf.tex
ms_caf submitted version

\shortauthors{Yu et al.} \shorttitle{CUBE}
\begin{document}
\title{CUBE: An \tcb{Information-optimized parallel} Cosmological $N$-body Algorithm}
\author{
Hao-Ran Yu$^{1,2,3}$,
Ue-Li Pen$^{2,4,5,6}$,
Xin Wang$^{2}$
}

@@

\begin{abstract}
Cosmological large scale structure $N$-body simulations are computation-light, memory-heavy problems in supercomputing. \tcb{The considerable amount of memory is usually dominated by an inefficient way of storing more than sufficient phase space information of particles.} We present a new \tcb{parallel, information-optimized}, particle-mesh based $N$-body code CUBE, \tcb{in which information- and memory-efficiency is increased by nearly an order of magnitude.} This is accomplished by storing particle's relative phase space coordinates \tcb{instead of global values, and in the format of fixed point as light as 1-byte. The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).} Our numerical experiments show that this \tcb{information-optimized $N$-body algorithm provides accurate results within the error of the particle-mesh algorithm.} This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.

@@

Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory to store the \tcb{information of particles}. Their phase space coordinates $(x,y,z,v_x,v_y,v_z)$ are stored as at least 6 single-precision floating point numbers (total of 24 bytes). On the other hand, modern supercomputer systems use multi cores, many integrated cores (MIC) and even densely parallelized GPUs, bringing orders of magnitude higher computing power, whereas these architectures usually have limited memory allocation. Thus, these computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are currently less suitable for fully usage of the computing power of modern supercomputers. For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power. \tcb{We investigate how most amount of information of particles can be deducted whereas it still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.}

In the following sections we present a new, \tcb{information-optimized} parallel cosmological $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle. It gives accurate results in cosmological LSS simulations -- \tcb{the error induced by information optimization is below the error from the PM algorithm}. In \S\ref{s.method} we show how the memory can be saved by using a ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format. In \S\ref{s.results} we quantify the accuracy of this algorithm by groups of simulations by {\tt CUBEP3M} and {\tt CUBE}. Discussions and conclusions are in \S\ref{s.discussion}.

\section{Method}\label{s.method}
The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle \tcb{(here after we use ``bpp'' referring to ``byte per particle'')}, which contains 4 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors. {\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}. This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh. Sometimes 4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations. If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid truncation errors. Another solution is to record {\it relative} coordinates for both position and velocity. {\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage with an integer based storage, reduces the basic memory usage from 28 bpp down to 6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}. The algorithm is described in \ref{ss.overview}.

@@

\begin{figure}
\centering
  \includegraphics[width=0.88\linewidth]{vdisp.pdf}
 \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}. The $x$-axis is the scale factor characterizing the cosmic evolution and the $y$-axis shows the $\sigma_v$ in unit of km/s. The 5 curves, from top to bottom, shows the $\sigma_v(a)$ with increasing smoothing scale $r$ from 0 to 20 Mpc$/h$.}
\label{fig.vdisp}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[width=0.90\linewidth]{fig_tile.pdf}
 \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy. In this example, there are 2 images per dimension (${\tt nn}=2$), and 2 tiles per image per dimension (${\tt nnt}=2$). The orange boxes show the overlapped physical+buffer=extended tile regions, inside of which one physical regions is indicated with green.}
\label{fig.tile}
\end{figure}

@@

\subsection{Code overview}\label{ss.overview}
{\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done. {\tt CUBE} is written in Coarray Fortran, where Coarray features \tcb{replace} MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran. We use this terminology in this paper.}. The algorithm is described in this language.

@@

{\tt timestep} --
We operate a Runge-Kutta 2 method for time integration. i.e., updating positions ($D$=drift) and velocities ($K$=kick) are at every half time step by \tcb{operator splitting}. For $n$ time steps, the operation would be $(DKKD)^n$ which is 2nd order accurate. The actual simulation applies varied time steps. In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, cosmic expansion and any other desired conditions.

@@

\begin{table}[]
\centering
\caption{Memory layout for a \tcb{certain} configuration}
\label{t.memory}
\begin{tabular}{llrrr}
\hline
& & \multicolumn{3}{c}{Memory usage}\\
\cline{3-5}
Type     & Array & /GB & /bpp & Percentage \\
\hline
Particles  & \tt xvp          & 29.9   & 8.24   & 83.8\%     \\
           & \tt xvp\_new     & \underline{3.16}   & \underline{0.872}  & \underline{8.87\%}     \\
           & \tt PID          & 0      & 0      &    0\%     \\
           & Subtotal         & 33.0   & 9.12   & 92.7\%     \\
\hline
Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.831\%    \\
           & \tt vfield       & 0.889  & 0.245  & 2.49\%     \\
           & \tt kern\_c      & 0.340  & 0.0939 & 0.954\%    \\
           & \tt force\_c     & (0.690)  & (0.190)  & (1.94\%)    \\
           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0394\% \\
           & Pencil-FFT       & (0.454)  & (0.125) & (1.27\%)      \\
           & Subtotal         & 1.55   & 0.429 & 4.36\%     \\
\hline
Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 2.97\%      \\
           & \tt force\_f     & (1.63) & (0.450) & (4.57\%)     \\
           & Fine-FFT     &   (1.41) & (0.389) & (3.96\%)     \\
           & Subtotal      & 1.06   & 0.292 & 2.97	\%      \\
\hline
Total &                      & 35.6 & 9.84 & 100\%\\
\hline
\multicolumn{2}{l} {Optimal limit} & 21.7 & 6 & 61.0\%\\
\hline
\end{tabular}
\end{table}

\begin{figure*}[]
\centering
  \includegraphics[width=0.75\linewidth]{overplotparticles.pdf}
 \caption{Offset in the particle distribution induced by integer-1 based algorithm (x1v1). In S512's subregion of volume $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$, particles from {\tt CUBEP3M} (bigger grey dots in the background) and {\tt CUBE}-x1v1 (smaller red dots) are projected onto the plane of $(15.625\ {\rm Mpc}/h)^2$. The comparing rules show 1 ${\rm Mpc}/h$, fine grid and coarse grid respectively, and position resolution of x1v1 is 1/64 of a fine grid.}
\label{fig.particles}
\end{figure*}

@@

\begin{table}[]
\centering
\caption{Simulation configurations}
\label{t.sim}
\begin{tabular}{lrrrrrr}
\hline
& \multicolumn{6}{c}{Configurations}\\
\cline{2-7}
Name  & $N_{\rm node}$ & $L/({\rm Mpc}/h)$ & $z_i$ & $z_{\rm ckpt}$ & $N_p$ & $m_p/\Msun$ \\
\hline
S512   & 8     & 200   & 49 & 0    & $512^3$  & $7.5\times 10^9$    \\
S256   & 1     & 80    & 49 & 0    & $256^3$  & $3.8\times 10^9$    \\
S2048S & 64    & 400   & 49 & 1,0  & $2048^3$ & $9.4\times 10^8$    \\
S2048L & 64    & 1200  & 49 & 1,0  & $2048^3$ & $2.5\times 10^{10}$ \\
\hline
TianNu & 13824 & 1200  & 100& 0.01 & $6912^3$ & $6.9\times 10^8$\\
       &       &       & 5  & 0.01 & $13824^3$& $3.2\times 10^5$\\
TianZero & 13824 & 1200 & 100& 0.01 & $6912^3$ & $7.0\times 10^8$\\
\hline
\end{tabular}
\end{table}

\section{Accuracy}\label{s.results}
We run a group of simulations to test the accuracy of {\tt CUBE}. We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively. Then the main $N$-body codes run their own initial conditions to redshift $z=0$. \tcb{We use the same force kernels as {\tt CUBEP3M} without PP force. Note that near find grid scales there are possibilities to enhance the force kernel to better match the nonlinear power spectrum predictions, however we use the conservative mode of {\tt CUBEP3M} in this paper. An extended PP force and an unbiased force matching algorithm will be added to {\tt CUBE}. The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, and here we focus on the cross correlations between different integer based methods.}

@@

We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass. These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE} with same initial seeds. Using different number of tiles per image gives exact same results. We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.

@@

\begin{figure}[]
\centering
  \includegraphics[width=0.99\linewidth]{ddsp_pcdf_256_box80.pdf}
 \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}. The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids. Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively. The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$. }
\label{fig.dsp}
\end{figure}

\begin{figure*}
\centering
  \includegraphics[width=1.0\linewidth]{ccc.pdf}
 \caption{\tcb{Fractional deviations $\Delta^2(k)/\Delta^2_0(k)-1$ in the power spectra relative to nonlinear predictions $\Delta^2_0(k)$ (upper axes), and {\tt CUBE}'s decorrelations $\xi(k)$ relative to {\tt CUBEP3M} (lower axes) induced by the integer-based algorithms. The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$. The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}. In lower panels, the dashed grey curves show the decorrelations given by the PM algorithm. The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$.}}
\label{fig.ccc}
\end{figure*}
\subsection{Power spectrum}

S2048S and S2048L are two simulations with $2048^3$ particles in small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes. We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$. The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}) to study halo properties. On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, baryonic acoustic oscillation (BAO) \citep{2005ApJ...633..560E} and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W} etc.

For each simulation, at redshifts $z=1$ and $z=0$, the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$. More generally we define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
\begin{equation}\label{eq.cps}
	\left\langle \delta_\alpha^\dagger({\bs k})\delta_\beta({\bs k}') \right\rangle=
    (2\pi)^3 P_{\alpha\beta}(k){\bs \delta}_{\rm 3D}({\bs k}-{\bs k}'),
\end{equation}
where ${\bs \delta}_{\rm 3D}$ is the three-dimensional Dirac delta function. In cosmology we usually consider the dimensionless power spectrum $\Delta^2_{\alpha\beta}(k)\equiv k^3 P_{\alpha\beta}(k)/(2\pi^2)$. The cross correlation coefficient is defined as
\begin{equation}\label{eq.ccc}
	\xi(k)\equiv P_{\alpha\beta}/\sqrt{P_{\alpha\alpha}P_{\beta\beta}}.
\end{equation}

\tcb{We compute the nonlinear prediction of the power spectrum by CLASS \citep{2011JCAP...07..034B}, denoted $\Delta^2_0(k)$, at $z=0,1$. Then we study the fractional power spectrum deviations $\Delta^2(k)/\Delta^2_0(k)-1$. In the four panels of figure \ref{fig.ccc} we denote the box sizes of the simulation and redshifts at the upper left corners. The upper parts of these four panels show $\Delta^2(k)/\Delta^2_0(k)-1$, and the four curves of different colors show the results of x2v2, x1v2, x2v1 and x1v1. They almost overlap with each other, and in general agree with the nonlinear predictions.}

\tcb{We label $k=0.2\,k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations. On this scale, the power spectra are offset from nonlinear predictions by at least 20\%. This error is from the PM algorithm and one has to increase the resolution of the simulation to correct these offsets. We do not plot {\tt CUBEP3M} because we examined that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results. Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.}

\tcb{In the lower parts of these four panels study the cross correlations. This time we in turn compare everything with {\tt CUBEP3M}, to emphasize the decorrelation by different integer formats. Firstly, x2v2 is perfectly correlated and cross correlation of x1v2 is nearly 100\% in all cases. In order to quantify the PM error in terms of decorrelation, we run x2v2 simulations with same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions. These results are dashed curves (labeled ``PM'') in the lower parts of these four panels. We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.}

\tcb{To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies. If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\,k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}

\section{Discussion and conclusion}\label{s.discussion}
We present an parallel, \tcb{information-optimized} $N$-body algorithm. This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g. \cite{2017PhRvD..95d3501Y,2017ApJ...841L..29W,2017MNRAS.469.1968P}. It needs very low memory usage, approaching 6 byte per particle (6 bpp).

The accuracy of this code is adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms. \tcb{Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies. In another word, the positions and velocities of each particle are exact. In practice, we only require that the errors from information optimization is much lower than the errors from PM algorithm. In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1. This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}

>>>>

\shortauthors{Yu et al.} \shorttitle{CUBE}
\begin{document}
\title{CUBE: An Information-optimized parallel Cosmological $N$-body Algorithm}
\author{
Hao-Ran Yu$^{1,2,3}$,
Ue-Li Pen$^{2,4,5,6}$,
Xin Wang$^{2}$
}

@@

\begin{abstract}
Cosmological large scale structure $N$-body simulations are computation-light, memory-heavy problems in supercomputing. The considerable amount of memory is usually dominated by an inefficient way of storing more than sufficient phase space information of particles. We present a new parallel, information-optimized, particle-mesh based $N$-body code CUBE, in which information- and memory-efficiency is increased by nearly an order of magnitude. This is accomplished by storing particle's relative phase space coordinates instead of global values, and in the format of fixed point as light as 1-byte. The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory). Our numerical experiments show that this information-optimized $N$-body algorithm provides accurate results within the error of the particle-mesh algorithm. This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.

@@

Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory to store the information of particles. Their phase space coordinates $(x,y,z,v_x,v_y,v_z)$ are stored as at least 6 single-precision floating point numbers (total of 24 bytes). On the other hand, modern supercomputer systems use multi cores, many integrated cores (MIC) and even densely parallelized GPUs, bringing orders of magnitude higher computing power, whereas these architectures usually have limited memory allocation. Thus, these computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are currently less suitable for fully usage of the computing power of modern supercomputers. For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power. We investigate how most amount of information of particles can be deducted whereas it still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.

In the following sections we present a new, information-optimized parallel cosmological $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle. It gives accurate results in cosmological LSS simulations -- the error induced by information optimization is below the error from the PM algorithm. In \S\ref{s.method} we show how the memory can be saved by using a ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format. In \S\ref{s.results} we quantify the accuracy of this algorithm by groups of simulations by {\tt CUBEP3M} and {\tt CUBE}. Discussions and conclusions are in \S\ref{s.discussion}.

\section{Method}\label{s.method}
The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle (here after we use ``bpp'' referring to ``byte per particle''), which contains 4 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors. {\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}. This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh. Sometimes 4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations. If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid truncation errors. Another solution is to record {\it relative} coordinates for both position and velocity. {\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage with an integer based storage, reduces the basic memory usage from 28 bpp down to 6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}. The algorithm is described in \ref{ss.overview}.

@@

\begin{figure}
\centering
  \includegraphics[width=1\linewidth]{f1}
 \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}. The $x$-axis is the scale factor characterizing the cosmic evolution and the $y$-axis shows the $\sigma_v$ in unit of km/s. The 5 curves, from top to bottom, shows the $\sigma_v(a)$ with increasing smoothing scale $r$ from 0 to 20 Mpc$/h$.}
\label{fig.vdisp}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[width=0.9\linewidth]{f2}
 \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy. In this example, there are 2 images per dimension (${\tt nn}=2$), and 2 tiles per image per dimension (${\tt nnt}=2$). The orange boxes show the overlapped physical+buffer=extended tile regions, inside of which one physical regions is indicated with green.}
\label{fig.tile}
\end{figure}

@@

\subsection{Code overview}\label{ss.overview}
{\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done. {\tt CUBE} is written in Coarray Fortran, where Coarray features replace MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran. We use this terminology in this paper.}. The algorithm is described in this language.

@@

{\tt timestep} --
We operate a Runge-Kutta 2 method for time integration. i.e., updating positions ($D$=drift) and velocities ($K$=kick) are at every half time step by operator splitting. For $n$ time steps, the operation would be $(DKKD)^n$ which is 2nd order accurate. The actual simulation applies varied time steps. In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, cosmic expansion and any other desired conditions.

@@

\begin{table}[]
\centering
\caption{Memory layout for a certain configuration}
\label{t.memory}
\begin{tabular}{llrrr}
\hline
& & \multicolumn{3}{c}{Memory usage}\\
\cline{3-5}
Type     & Array & /GB & /bpp & Percentage \\
\hline
Particles  & \tt xvp          & 29.9   & 8.24   & 83.8\%     \\
           & \tt xvp\_new     & \underline{3.16}   & \underline{0.872}  & \underline{8.87\%}     \\
           & \tt PID          & 0      & 0      &    0\%     \\
           & Subtotal         & 33.0   & 9.12   & 92.7\%     \\
\hline
Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.831\%    \\
           & \tt vfield       & 0.889  & 0.245  & 2.49\%     \\
           & \tt kern\_c      & 0.340  & 0.0939 & 0.954\%    \\
           & \tt force\_c     & (0.690)  & (0.190)  & (1.94\%)    \\
           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0394\% \\
           & Pencil-FFT       & (0.454)  & (0.125) & (1.27\%)      \\
           & Subtotal         & 1.55   & 0.429 & 4.36\%     \\
\hline
Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 2.97\%      \\
           & \tt force\_f     & (1.63) & (0.450) & (4.57\%)     \\
           & Fine-FFT     &   (1.41) & (0.389) & (3.96\%)     \\
           & Subtotal      & 1.06   & 0.292 & 2.97	\%      \\
\hline
Total &                      & 35.6 & 9.84 & 100\%\\
\hline
\multicolumn{2}{l} {Optimal limit} & 21.7 & 6 & 61.0\%\\
\hline
\end{tabular}
\end{table}

\begin{figure*}[]
\centering
  \includegraphics[width=0.75\linewidth]{f7}
 \caption{Offset in the particle distribution induced by integer-1 based algorithm (x1v1). In S512's subregion of volume $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$, particles from {\tt CUBEP3M} (bigger grey dots in the background) and {\tt CUBE}-x1v1 (smaller red dots) are projected onto the plane of $(15.625\ {\rm Mpc}/h)^2$. The comparing rules show 1 ${\rm Mpc}/h$, fine grid and coarse grid respectively, and position resolution of x1v1 is 1/64 of a fine grid.}
\label{fig.particles}
\end{figure*}

@@

\begin{table}[]
\centering
\caption{Simulation configurations}
\label{t.sim}
\begin{tabular}{lrrrrr}
\hline
& \multicolumn{5}{c}{Configurations}\\
\cline{2-6}
Name  & $N_{\rm node}$ & $L/({\rm Mpc\,}h^{-1})$ & $z_i$ & $N_p$ & $m_p/\Msun$ \\
\hline
S512   & 8     & 200   & 49   & $512^3$  & $7.5\times 10^9$    \\
S256   & 1     & 80    & 49   & $256^3$  & $3.8\times 10^9$    \\
S2048S & 64    & 400   & 49   & $2048^3$ & $9.4\times 10^8$    \\
S2048L & 64    & 1200  & 49   & $2048^3$ & $2.5\times 10^{10}$ \\
\hline
TianNu & 13824 & 1200  & 100  & $6912^3$ & $6.9\times 10^8$\\
       &       &       & 5    & $13824^3$& $3.2\times 10^5$\\
TianZero & 13824 & 1200 & 100 & $6912^3$ & $7.0\times 10^8$\\
\hline
\end{tabular}
\end{table}

\begin{figure}[]
\centering
  \includegraphics[width=0.99\linewidth]{f8}
 \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}. The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids. Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively. The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$. }
\label{fig.dsp}
\end{figure}

\section{Accuracy}\label{s.results}
We run a group of simulations to test the accuracy of {\tt CUBE}. We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively. Then the main $N$-body codes run their own initial conditions to redshift $z=0$. We use the same force kernels as {\tt CUBEP3M} without PP force. Note that near find grid scales there are possibilities to enhance the force kernel to better match the nonlinear power spectrum predictions, however we use the conservative mode of {\tt CUBEP3M} in this paper. An extended PP force and an unbiased force matching algorithm will be added to {\tt CUBE}. The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, and here we focus on the cross correlations between different integer based methods.

@@

We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, total number of particles and particle mass. These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE} with same initial seeds. Using different number of tiles per image gives exact same results. We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.

\begin{figure*}
\centering
  \includegraphics[width=1.0\linewidth]{f9}
 \caption{Dimensionless power spectra $\Delta^2(k)$ and decorrelations $\xi(k)$ (lower axes) induced by the integer-based algorithms in simulations S2048S and S2048L. The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}. In lower panels, the dashed curves show the decorrelations given by the PM algorithm, measured by cross correlating different resolutions of {\tt CUBE}-x2v2. The vertical orange dashed lines show the scale $k=0.2\, k_{\rm Nyquist}$.}
\label{fig.ccc}
\end{figure*}
\subsection{Power spectrum}

@@

S2048S and S2048L are two simulations with $2048^3$ particles in small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes. We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at $z=0$. The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}). On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, baryonic acoustic oscillation (BAO) \citep{2005ApJ...633..560E} and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W} etc.

For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$. We define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
\begin{equation}\label{eq.cps}
	\left\langle \delta_\alpha^\dagger({\bs k})\delta_\beta({\bs k}') \right\rangle=
    (2\pi)^3 P_{\alpha\beta}(k){\bs \delta}_{\rm 3D}({\bs k}-{\bs k}'),
\end{equation}
where ${\bs \delta}_{\rm 3D}$ is the three-dimensional Dirac delta function. In cosmology we usually consider the dimensionless power spectrum $\Delta^2_{\alpha\beta}(k)\equiv k^3 P_{\alpha\beta}(k)/(2\pi^2)$. The cross correlation coefficient is defined as
\begin{equation}\label{eq.ccc}
	\xi(k)\equiv P_{\alpha\beta}/\sqrt{P_{\alpha\alpha}P_{\beta\beta}}.
\end{equation}

In the upper two panels of figure \ref{fig.ccc} we show the power spectra of {\tt CUBE}. In both plots of S2048S and S2048L the four solid curves of different colors show the results of x2v2, x1v2, x2v1 and x1v1, and they almost overlapped with each other. The dashed curves are the nonlinear prediction of matter power spectrum by CLASS \citep{2011JCAP...07..034B}.

We label $k=0.2\,k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations. On this scale, the power spectra are offset from nonlinear predictions by at least 20\%. This error is from the PM algorithm and one has to increase the resolution of the simulation to correct these offsets. We do not plot {\tt CUBEP3M} because we examined that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results. Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.

In the lower parts of these four panels study the cross correlations. We compare everything with {\tt CUBE}-x2v2, to emphasize the decorrelation by different integer formats. Note that x2v2 is perfectly correlated with {\tt CUBEP3M}. In the lower two panels of figure \ref{fig.ccc} the solid curves show the decorrelation $\xi$ of x1v2, x2v1 and x1v1. Higher resolution (S2048S) in general comes with more decorrelations, and x1v2 cross correlate with x2v2 better than the other two cases. In order to quantify the PM error in terms of decorrelations, we run x2v2 simulations with same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions. These results are shown in dashed curves (labeled ``PM''). We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.

To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies.

\section{Discussion and conclusion}\label{s.discussion}
We present an parallel, information-optimized $N$-body algorithm. This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g. \cite{2017PhRvD..95d3501Y,2017ApJ...841L..29W,2017MNRAS.469.1968P}. It needs very low memory usage, approaching 6 byte per particle (6 bpp).

The accuracy of this code is adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms. Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies. In another word, the positions and velocities of each particle are exact. In practice, we only require that the errors from information optimization is much lower than the errors from PM algorithm. In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1. This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.
